---
title: "Lab Assignment 5"
author: "Michael Beebe"
date: "9/21/2020"
output: html_document
---  

```{r Setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
```

Loading the Old Faithful dataset
```{r Loading the Old Faithful dataset}
data("faithful")
head(faithful)
```

2.1. Histogram
--
```{r Base R histogram}
hist(
  x = faithful$eruptions,
  main = "Old Faithful Eruptions",
  xlab = "Eruption Times (mins)",
  ylab = "Time Between Eruptions (mins)"
)
```

2.2. Base Scatterplot
--
```{r Base R scatterplot 1}
scatter.smooth(
  x = faithful$eruptions,
  y = faithful$waiting,
  main = "Old Faithful Eruptions",
  xlab = "Eruption Times (mins)",
  ylab = "Time Between Eruptions (mins)"
)
```

```{r Base R scatterplot 2}
scatter.smooth(
  x = faithful$waiting,
  y = faithful$eruptions,
  main = "Old Faithful Eruptions",
  xlab = "Time Between Eruptions (mins)",
  ylab = "Eruption Times (mins)"
)
```

According to these scatterplots, longer eruptions have longer waiting times to the next eruption, and shorter eruptions have shorter waiting times in between eruptions.  

2.3. Boxplots
--
```{r Base R boxplot}
boxplot(
  x = faithful
)
```
From this boxplot, we can see that the wait times are much longer than the eruptions themselves.  

3. ggplot2
--

3.1. Histograms and Densities
--
```{r Histograms}
# ggplot2 loaded at the beginning of the md
ggplot(data = faithful, aes(x=eruptions)) + geom_histogram(binwidth = 1)
```

```{r Densities}
ggplot(data = faithful, aes(x=eruptions)) + geom_density(color = "blue")
```

3.2. Scatterplots
--
```{r Scatterplots}
f <- 22
ggplot(data = faithful, aes(x = eruptions, y = waiting, fill = "blue")) + geom_point(shape = f)
```

3.3. Boxplots
--
```{r Boxplots}
ggplot(data = faithful, aes(x=eruptions)) + geom_boxplot()
```

3.4. Line Graphs
--
```{r Line Graphs}
ggplot(data = faithful, aes(x=eruptions, y = waiting)) + geom_line()
```

