---
title: 'Lab Assignment: Chapter 17 - Probablity Distributions'
author: "Michael Beebe"
output:
  word_document: default
  html_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Dependencies}
library(ggplot2)
library(reshape2)
library(stringr)
```

1. Probability Distribution
--

Using rnorm store 10 random numbers in a variable. 
```{r rnorm}
rand10 <- rnorm(10)
rand10
```

Find the probabilities of your numbers from above.  
```{r probabilities}
rand10Probs <- dnorm(rand10)
rand10Probs
```

Create 2 new variables - one that will store a large number of random variables, and the other that stores their density (probabilities)
```{r Create 2 new variables}
rand1000 <- rnorm(1000)
rand1000Probs <- dnorm(rand1000)
```

Plot the distribution using ggplot.
```{r Plot the distribution using ggplot.}
ggplot(
  data.frame(x=rand1000, y=rand1000Probs)) + 
  aes(x=x, y=y) +geom_point() + 
  labs(x="Random Normal Variables", y="Density"
)
```

2. Binomial Distribution
--

Calculate the probability and cumulative probability using the Binomial Distribution using the following scenarios: 

Probability of tossing a die 10 times and getting a "3" at least 2 times.  

* Probability of a success is 1/6 = 0.17  
```{r dbinom}
dbinom(x = 2, size = 10, prob = .17)
```

Probability and getting at least a "B" (80%) on a quiz that has 10 true/false questions.  

* Probability of getting a question right is 1/2 = 0.5.  
* Use dbinom function to determine probability of getting a B by randomly guessing  
```{r Use dbinom function to determine probability of getting a B by randomly guessing.}
dbinom(x = 8, size = 10, prob = .5)
```

* Use pbinom function to determine probability of answering 8 or less questions correctly by random guess.  
```{r Use pbinom function to determine probability of answering 8 or less questions correctly by random guess.}
pbinom(q = 8, size = 10, prob = .5)
```

Generate a graph based on a 10,000 random experiments, 10 trials, and 0.3 probability of success.  
```{r Generate a graph based on a 10,000 random experiments, 10 trials, and 0.3 probability of success.}
binomData <- data.frame (Successes= rbinom (n=10000, size=10, prob=.3))
ggplot (data = binomData, aes (x=Successes)) + geom_histogram (binwidth=1)
```

3. Poisson Distribution
--

We are interested in calculating the probability that today at least 5 callers will receive a busy signal if they call Lakeland's Help Desk.  

* The average probability of a busy signal is 0.5%  
* There are 1,200 calls to the Lakeland Help Desk  
* Use the ppois function with lower=FALSE
```{r Help Desk}
ppois(q = 1200, lambda = 0.005, lower.tail = FALSE) 
```

Generate the plot on page 238 and explain what it shows.
```{r pg 238 plot}
pois1 <- rpois(n = 10000, lambda = 1)
pois2 <- rpois(n = 10000, lambda = 2)
pois5 <- rpois(n = 10000, lambda = 5)
pois10 <- rpois(n = 10000, lambda = 10)
pois20 <- rpois(n = 10000, lambda = 20)
pois <- data.frame(
  Lambda.1 = pois1,
  Lambda.2 = pois2,
  Lambda.5 = pois5,
  Lambda.10 = pois10,
  Lambda.20 = pois20
)

pois <- melt(data=pois, variable.name="Lambda", value.name="x")
pois$Lambda <- as.factor(
  as.numeric(
    str_extract(string=pois$Lambda, pattern="\\d+")
  )
)


ggplot (pois, aes (x=x)) +
geom_density ( aes (group=Lambda, color=Lambda, fill=Lambda), adjust=4, alpha=1/2) +
scale_color_discrete () + scale_fill_discrete () +
ggtitle ("Probability Mass Function")
```

This graph shows density plots for 10,000 draws from the Poisson distribution at varying levels of λ.