---
title: "Chapter 12 - Faster Group Manipulation with dplyr"
author: "Michael Beebe"
output:
  html_document: default
  word_document: default
  pdf_document: default
---  
--------------------------------------------------------------------------------    
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(readr)
```

1. dplyr  
--  

From the textbook data files, load the the housing.csv data
```{r From the textbook data files, load the the housing.csv data}
housing <- read.csv("../../Data/housing.csv")
head(housing, 5)
```

2. Select  
--  

Using the housing data and basic select, display the neighborhood and boro
```{r Using the housing data and basic select, display the neighborhood and boro}
neighborhood_boro <-
  housing %>%
    select(Neighborhood, Boro)

head(neighborhood_boro, 10)
```

Select neighborhood and year built by using a vector passed to the select function
```{r Select neighborhood and year built by using a vector passed to the select function}
neighborhood_year <-
  housing %>%
    select(c(Neighborhood, Year.Built))

head(neighborhood_year, 10)
# I am not getting a warning
```

Select the neighborhood and square footage using the column positions rather than names
```{r Select the neighborhood and square footage using the column positions rather than names}
neighborhood_sqft <-
  housing %>%
    select(1, 5)

head(neighborhood_sqft, 10)
```

Select the neighborhood and the columns that start with E
```{r Select the neighborhood and the columns that start with E}
neighborhood_e_cols <-
  housing %>%
    select(Neighborhood, starts_with("e"))

head(neighborhood_e_cols, 10)
```

Select the neighborhood and the columns that end with t
```{r Select the neighborhood and the columns that end with t}
neighborhood_t_cols <-
  housing %>%
    select(Neighborhood, ends_with("t"))

head(neighborhood_t_cols, 10)
```


Select the columns that match the pattern ".Income". The word Income should be at the end of the string.
```{r Select the columns that match the pattern ".Income". The word Income should be at the end of the string.}
neighbord_income <-
  housing %>%
    select(Neighborhood, ends_with(".income"))

head(neighbord_income, 10)
```

3. Filter
--

Using the housing data, filter the data based on construction built after or equal to the year 2009
```{r Using the housing data, filter the data based on construction built after or equal to the year 2009}
after_2009 <-
  housing %>%
    filter(Year.Built >= 2009)

head(after_2009, 5)
```

Issue a select and store the Neighborhood, Year Built, and Boro in a variable
```{r Issue a select and store the Neighborhood, Year Built, and Boro in a variable, message=FALSE}
neighborhood_yb_boro <-
  housing %>%
    select(Neighborhood, Year.Built, Boro)

head(neighborhood_yb_boro, 5)
```

Filter on Boro in the Bronx or Brooklyn
```{r Filter on Boro in the Bronx or Brooklyn}
bronx_brooklyn <-
  neighborhood_yb_boro %>%
    filter(Boro == "Bronx" | Boro == "Brooklyn")

head(bronx_brooklyn, 5)
tail(bronx_brooklyn, 5)
```

Using the housing data, filter on Year Built > 1999 and Total Units > 200
```{r Using the housing data, filter on Year Built > 1999 and Total Units > 200}
yb_tu <-
  housing %>%
    filter(Year.Built > 1999 & Total.Units > 200)

head(yb_tu, 5)
```

Declare 2 variables - theCol and theVal  
 - Use filter and sprintf to filter the housing data on Neighborhoods in the Financial District  
 - Disregard warning message if you receive one
```{r Declare 2 variables - theCol and theVal }
theCol <- housing$Neighborhood
theVal <- sprintf("%s", "FINANCIAL")

fincancial_district <-
  housing %>%
    filter(theCol == theVal)

head(fincancial_district, 5)
```

4. Slice
--  

Using the housing data, take a slice of rows 10-20
```{r Using the housing data, take a slice of rows 10-20}
housing %>%
  slice(10:20)
```

Now, using the same data take a slice of rows 1-5 and the last 5 rows 
```{r Now, using the same data take a slice of rows 1-5 and the last 5 rows}
housing %>%
  slice(c(1:5, (n()-4):n()))
```

5. Mutate
--  
Create a new column called Age  
 - This column will subtract the year built from the current year  
 - You will need to pipe select and mutate
```{r Create a new column called Age}
Age <-
  housing %>%
    select(Neighborhood, Year.Built) %>%
    mutate(Age = 2020 - Year.Built)

head(Age)
```

6. Summarize and Group By
--  

Using the housing date and the summarize function, find the mean square footage
```{r Using the housing date and the summarize function, find the mean square footage}
mean_sqft <-
  housing %>%
    summarize("Mean Sq Ft" = mean(Gross.SqFt))

mean_sqft
```

Using summarize and group by, find the mean square footage and group by Neighborhood
```{r Using summarize and group by, find the mean square footage and group by Neighborhood, message=FALSE}
mean_sqft_neighborhood <-
  housing %>%
    group_by(Neighborhood) %>%
    summarize("Mean Sq Ft" = mean(Gross.SqFt))

head(mean_sqft_neighborhood, 5)
```
