---
title: "4. Matrices"
author: "Michael Beebe"
date: "9/9/2020"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

```

```{r Create a matrix called mymat1}

mymat1 <- matrix(
  data = c(4.3, 3.1, 8.2, 8.2, 3.2, 0.9, 1.6, 6.5),
  nrow = 4,
  ncol = 2,
  byrow = FALSE
)

```

```{r Create a matrix called mymat2}

mymat2 <- matrix(
  data = c(4.3, 3.1, 8.2, 8.2, 3.2, 0.9, 1.6, 6.5),
  nrow = 4,
  ncol = 2,
  byrow = FALSE
)

```

```{r Include code that will print each matrix to the console}

print(mymat1)
print(mymat2)

```

```{r Display the number of rows and number of columns to the console}

glue::glue("Number of rows in mymat1: {nrow(mymat1)}")
glue::glue("Number of columns in mymat1: {ncol(mymat1)}")

glue::glue("Number of rows in mymat2: {nrow(mymat2)}")
glue::glue("Number of columns in mymat2: {ncol(mymat2)}")

```

```{r Add mymat1 and mymat2}

mymat3 <- mymat1 + mymat2
mymat3

```

```{r Multiply mymat1 and mymat2}

mymat4 <- mymat1 * mymat2
mymat4

```

```{r Check to see if the elements are equal in each matrix}

mymat1
mymat2

```

The values of mymat1 and mymat2 are still equal because I assigned the added/multiplied values to a new matrix. In order to change the values of mymat1 or mymat2, we would need to reassign their value to mymat1 + mymat2 or mymat1 * mymat2.

```{r Perform matrix multiplication by transposing mymat2 onto mymat1}

mymat5 <- t(mymat2) * t(mymat1)
mymat5

```

Matrix multiplication is a commonly used operation in mathematics, requiring the number of columns of the left-hand matrix to be the same as the number of rows of the right-hand matrix. 

