---
title: 'Lab Assignment: Chapter 18 - Basic Statistics'
author: "Michael Beebe"
output:
  word_document: default
  html_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Dependencies, warning=FALSE}
library(ggplot2)
library(GGally)
library(MASS)
library(reshape2)
library(plyr)
library(dplyr)
```

1. Summary Stastics
--  

Create a variable - stats - and store a random sample of 100 numbers between 100 and 200. Use replace=TRUE 
```{r}
stats <- sample(x = 100:200, size = 100, replace = TRUE)
head(stats)
```

Calculate the mean
```{r}
mean(stats)
```

Calculate the variance (2 ways)
```{r}
sum((stats - mean(stats))^2) / (length(stats) - 1)
```

```{r}
var(stats)
```

Calculate the standard deviation (2 ways)
```{r}
sqrt(var(stats))
```

```{r}
sd(stats)
```


Display the summary statistics
```{r}
summary(stats)
```

Display the 1st and 3rd quartiles in a different way than the summary statistics
```{r}
quantile(x = stats, probs = c(.25, .75))
```

A student's grades are the following: quiz 1= 78, quiz 2 = 74, quiz 3 = 78, quiz 4 = 76, midterm = 68, and final = 90. The quizzes are weighted at 1/8 of the final grade and the exams are weighted at 1/4.  
```{r}
grades <- c(78, 74, 78, 76, 68, 90)
weights <- c(1/8, 1/8, 1/8, 1/8, 1/4, 1/4)
```

* Display the mean.
```{r}
mean(grades)
```

Display the weighted mean.
```{r}
weighted.mean(x = grades, w = weights)
```

2. Correlation and Covariance
--

Using the Old Faithful data, store the waiting time in a variable and the eruption duration in another
```{r}
wait <- faithful$waiting
duration <- faithful$eruptions
```

Find the correlation coefficient of eruption duration and waiting time in the data set faithful.
```{r}
cor(wait, duration)
```

```{r}
cor(faithful)
```

Observe if there is any linear relationship between the variables.  
```{r}
ggplot(data.frame(x=wait, y=duration), aes(x=x, y=y)) +
  geom_point() +
  labs(x ="Wait Time", y = "Eruption Time", title = "Old Faithful")
```


* We can see that there is very much so a linear correlation between wait time between eruptions and the duration of the eruption.

Using the same variables find the covariance, and analyze the results
```{r}
cov(wait, duration)
```

```{r}
cov(faithful)
```

Use the cor function on a matrix to evaluate the faithful data.
```{r}
mat <- cbind(wait, duration)
cor(mat)
```

Graph the Old Faithful data
```{r}
ggpairs(faithful)
```

3. T-Tests
--  

Use the built-in data set named immer, which contains data on the barley yield in years 1931 and 1932 of a field.  
```{r}
head(immer)
```

The data follows the normal distribution. Find the 95% confidence interval estimate of the difference between the mean barley yields.  
```{r}
t.test(immer$Y1, immer$Y2, paired = TRUE, conf.level = 0.95)
```

Analyze the results  
```{r}
yieldDiff <- immer$Y2 - immer$Y1

ggplot(immer, aes(x=Y2 - Y1)) +
  geom_density() +
  geom_vline(xintercept = mean(yieldDiff)) +
  geom_vline(xintercept = mean(yieldDiff) + 2*c(-1,1) * sd(yieldDiff) / sqrt(nrow(immer)), linetype=2)
```


4. ANOVA
--  

Complete the Analysis of Variance for the TIPS data
```{r}
data(tips, package = "reshape2")
head(tips)
```

```{r message=FALSE}
ggpairs(tips)
```


```{r}
tipAnova <- aov(tip ~ day - 1, tips)
tipIntercept <- aov(tip ~ day, tips)
```

```{r}
tipAnova$coefficients
```

```{r}
tipIntercept$coefficients
```

```{r}
summary(tipAnova)
```

Analyze the p-values for each set of summary data.  
```{r}
tipsByDay <- ddply(
  tips,
  "day",
  summarize,
  tip.mean=mean(tip),
  Length=NROW(tip),
  tip.sd=sd(tip),
  tfrac=qt(p=.90, df=Length-1),
  Lower=tip.mean - tfrac*tip.sd/sqrt(Length),
  Upper=tip.mean + tfrac*tip.sd/sqrt(Length)
)
tipsByDay
```

Plot the dataset.
```{r}
ggplot(tipsByDay, aes(x=tip.mean, y=day)) +
  geom_point() +
  geom_errorbarh(aes(xmin=Lower, xmax=Upper), height=.3) +
  labs(x="Mean Tips", y="Day", title = {"Mean Tips by Day"})
```

Complete the Analysis of Variance for the Old Faithful data.
```{r}
head(faithful)
```

```{r}
faithfulAnova <- aov(eruptions ~ waiting -1, faithful)
faithfulIntercept <- aov(eruptions ~ waiting, faithful)
```

```{r}
faithfulAnova$coefficients
```

```{r}
faithfulIntercept$coefficients
```

```{r}
summary(faithfulAnova)
```

Analyze the p-values for each set of summary data.  
```{r warning=FALSE}
eruptionsByWait <- ddply(
  faithful,
  "waiting",
  summarize,
  eruptions.mean=mean(eruptions),
  Length=NROW(eruptions),
  eruptions.sd=sd(eruptions),
  tfrac=qt(p=.90, df=Length-1),
  Lower=eruptions.mean - tfrac*eruptions.sd/sqrt(Length),
  Upper=eruptions.mean + tfrac*eruptions.sd/sqrt(Length)
)
head(eruptionsByWait)
```

Plot the dataset.
```{r message=FALSE}
ggpairs(faithful)
```