---
title: 'Lab Assignment: Chapter 15 - Reshaping Data in the Tidyverse'
author: "Michael Beebe"
output:
  word_document: default
  html_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Dependencies}
library(dplyr)
library(tibble)
library(readr)
library(tidyr)
```

1. Binding Rows and Columns
--

Create 2 tibbles:

* tbl1 will have the numbers 1-5 as IDs, and the first names of 5 characters from your favorite TV show  
* tbl2 will have the numbers 1-4, & 6 as IDs, and the last names of the characters 1-4 above, and the last name of a different character that doesn't match from tbl1
```{r Create 2 tibbles}
tibble1 <- tibble(
  ID = c(1:5),
  F_Names = c("Michael", "Dwight", "Jim", "Pam", "Toby")
)

tibble2 <- tibble(
  ID = c(1:4, 6),
  L_Names = c("Scott", "Schrute", "Halpert", "Beasley", "Malone")
)

tibble1
tibble2
```

Join the rows and print the results to the screen
```{r Join the rows and print the results to the screen}
tibble3 <- full_join(tibble1, tibble2)

tibble3
```

2. Joins with dplyr
--

Add a line or lines to your script that will join the tables from step 1 above into 1 table  

* Use the ID field to join the tables  
* Use a left join and a take a screen shot of your results  
* Use a right join and take a screen shot of your results
```{r Left join}
left <- left_join(
  tibble1,
  tibble2,
  by = "ID"
)

left
```

```{r Right join}
right <- right_join(
  tibble1,
  tibble2,
  by = "ID"
)

right
```


3. Converting Data Formats
--  

Use the read_tsv file function to read the "reaction.txt" file from the book site  

* Store the data into a variable "emotion"  
* Print emotion to the screen
```{r Use the read_tsv file function to read the "reaction.txt" file from the book site  }
emotion <- read_tsv("../../Data/reaction.txt")
head(emotion)
```

Gather the data with a key = Type and value = Measurement
```{r Gather the data}
emotion %>%
  gather(
    key = Type,
    value = Measurement
  )
```

Arrange the data from the step above by ID
```{r Arrange the data from the step above by ID}
emotion %>%
  arrange(ID)
```




