---
title: "ITCS 1880: Final Project"
author: "By Michael Beebe"
output:
  html_document: default
  word_document: default
---

-------------------------------------------------

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loading in dependencies
```{r Dependencies, message=FALSE}
library(tidyverse)
library(glue)
library(ggplot2)
library(hrbrthemes)
library(psych)
library(graphics)
```

Loading in dataset to be analyzed
```{r}
customer_data <- read.csv("Data/Mall_Customers.csv")
```


# Exploratory Data Analysis  

The number of rows and the number of columns in the data set.
```{r}
nrow(customer_data)
```

```{r}
ncol(customer_data)
```

The names of the columns in the data set
```{r}
colnames(customer_data)
```


The top 10 rows of the data set
```{r}
head(customer_data, 10)
```

The class of the dataset and each column
```{r}
class(customer_data)
```
```{r}
lapply(customer_data, class)
```

Summarize the data in the Mall Customers data set. Be sure to include: Min, Max, Mean, Median, Quartiles and Standard Deviation:  

### Age

Min
```{r}
glue("Min Age: {min(customer_data$Age)}")
```

Max   
```{r}
glue("Max Age: {max(customer_data$Age)}")
```
  
Mean    
```{r}
glue("Mean Age: {mean(customer_data$Age)}")
```
  
Median
```{r}
glue("Median Age: {median(customer_data$Age)}")
```
  
Quartiles
```{r}
quantile(customer_data$Age)
```
  
Standard Deviation
```{r}
glue("Standard Deviation of Age: {sd(customer_data$Age)}")
```

```{r}
describe(customer_data)
describe.by(customer_data, group = customer_data$Gender)
```

### Annual Income ($1,000's)

Min
```{r}
glue("Min Annual Income in $1,000's: {min(customer_data$Annual.Income..k..)}")
```

Max
```{r}
glue("Max Annual Income in $1,000's: {max(customer_data$Annual.Income..k..)}")
```
  
Mean
```{r}
glue("Mean Annual Income in $1,000's: {mean(customer_data$Annual.Income..k..)}")
```
  
Median
```{r}
glue("Median Annual Income in $1,000's: {median(customer_data$Annual.Income..k..)}")
```
  
Quartiles
```{r}
quantile(customer_data$Annual.Income..k..)
```

Standard Deviation
```{r}
glue("Standard Deviation of Annual Income in $1,000's: {sd(customer_data$Annual.Income..k..)}")
```

### Spending Score (1-100)

Min
```{r}
glue("Min Spending Score: {min(customer_data$Spending.Score..1.100.)}")
```

Max
```{r}
glue("Max Spendinc Score: {max(customer_data$Spending.Score..1.100.)}")
```
  
Mean
```{r}
glue("Mean Spending Score: {mean(customer_data$Spending.Score..1.100.)}")
```
  
Median
```{r}
glue("Median Spending Score: {median(customer_data$Spending.Score..1.100.)}")
```
  
Quartiles
```{r}
quantile(customer_data$Spending.Score..1.100.)
```

Standard Deviation
```{r}
glue("Standard Deviation of Spending Score: {sd(customer_data$Spending.Score..1.100.)}")
```
### Observations by Gender

Average age by gender
```{r}
age_by_gender <- aggregate(
  x = customer_data$Age,
  by = list(customer_data$Gender),
  FUN = mean
)
age_by_gender
```

Average annual income by gender
```{r}
income_by_gender <- aggregate(
  x = customer_data$Annual.Income..k..,
  by = list(customer_data$Gender),
  FUN = mean
)
income_by_gender
```

Average spending score by gender
```{r}
spending_by_gender <- aggregate(
  x = customer_data$Spending.Score..1.100.,
  by = list(customer_data$Gender),
  FUN = mean
)
spending_by_gender
```

New DF combining the three above aggregations
```{r}
summarized_avgs_by_gender <- data.frame(
  Gender = c("Female", "Male"),
  Average_Age = age_by_gender$x,
  Average_Annual_Income = income_by_gender$x,
  Avergae_Spending = spending_by_gender$x
)
summarized_avgs_by_gender
```

### Data Analysis Summary
!!!!!!!!  TODO  !!!!!!!!

-------------------------------------------------------------------

# Plotting

### Boxplots

```{r}
ggplot(data = customer_data, aes(x = Age)) +
  geom_boxplot() +
  ggtitle("Age Distribution and Variability") +
  xlab("Age") +
  theme_get()
```

```{r}
ggplot(data = customer_data, aes(x = Annual.Income..k..)) +
  geom_boxplot() +
  ggtitle("Income Distribution and Variability") +
  xlab("Annual Income ($1,000's)") +
  theme_get()
```

```{r}
ggplot(data = customer_data, aes(x = Spending.Score..1.100.)) +
  geom_boxplot() +
  ggtitle("Spending Score Distribution and Variability") +
  xlab("Spending Score (1-100)") +
  theme_get()
```

### Histograms (By Gender)

```{r}
ggplot(data = customer_data, mapping = aes(x=Gender)) +
  geom_bar() +
  ggtitle ("Number of Observations") +
  xlab("Gender") +
  ylab("Count") +
  theme_get()
```

```{r message=FALSE}
ggplot(data = summarized_avgs_by_gender, mapping = aes(x=Gender, y=Average_Age)) +
  geom_histogram(stat="identity") +
  ggtitle ("Average Age by Gender") +
  xlab("Gender") +
  ylab("Average Age") +
  theme_get()
```

```{r}
ggplot(data = summarized_avgs_by_gender, mapping = aes(x=Gender, y=Average_Annual_Income)) +
  geom_histogram(stat="identity") +
  ggtitle ("Avg Annual Income by Gender") +
  xlab("Gender") +
  ylab("Avg Annual Income ($1,000's)") +
  theme_get()
```

```{r}
ggplot(data = summarized_avgs_by_gender, mapping = aes(x=Gender, y=Avergae_Spending)) +
  geom_histogram(stat="identity") +
  ggtitle ("Avg Spending Score by Gender") +
  xlab("Gender") +
  ylab("Avg Spending Score") +
  theme_get()
```

### Histograms (By Age Group)

```{r}
#         ADDING AN AGE GROUP COLUMN TO THE CUSTOMER DATA DF
age_cut <- cut(customer_data$Age, breaks = c(18, 30, 40, 50, 60, 70, 80), right = FALSE)
customer_data_2 <- data.frame(customer_data, Age_Group = age_cut)
head(customer_data_2)
```

```{r}
ggplot(customer_data_2, aes(x = Age_Group)) +
  geom_bar() +
  ggtitle ("Number of Observations by Age") +
  xlab("Age Group") +
  ylab("Count") +
  theme_get()
```

```{r}
income_by_age_group <- aggregate(
  x = customer_data_2$Annual.Income..k..,
  by = list(customer_data_2$Age_Group),
  FUN = mean
)
income_by_age_group
```

```{r}
spending_by_age_group <- aggregate(
  x = customer_data_2$Spending.Score..1.100.,
  by = list(customer_data_2$Age_Group),
  FUN = mean
)
spending_by_age_group
```

```{r}
summarized_avgs_by_age_group <- data.frame(
  Age_Group = income_by_age_group$Group.1,
  Avg_Income = income_by_age_group$x,
  Avg_Spending = spending_by_age_group$x
)
summarized_avgs_by_age_group
```

```{r}
ggplot(data = summarized_avgs_by_age_group, mapping = aes(x=Age_Group, y=Avg_Income)) +
  geom_histogram(stat="identity") +
  ggtitle ("Avg Annual Income by Age Group") +
  xlab("Age Group") +
  ylab("Avg Annual Income ($1,000's)") +
  theme_get()
```

```{r}
ggplot(data = summarized_avgs_by_age_group, mapping = aes(x=Age_Group, y=Avg_Spending)) +
  geom_histogram(stat="identity") +
  ggtitle ("Avg Spending by Age Group") +
  xlab("Age Group") +
  ylab("Avg Spending Score (1-100)") +
  theme_get()
```

### Scatterplot
```{r}
ggplot(data = customer_data, mapping = aes(x=Annual.Income..k.., y=Spending.Score..1.100.)) +
  geom_point() +
  ggtitle("Income and Spending (Whole Dataset)") +
  xlab("Annual Income ($1,000)") +
  ylab("Spending Score (1-100)") +
  theme_get()
```

```{r}
male_df <- customer_data_2[customer_data_2$Gender == "Male", ]
head(male_df)

female_df <- customer_data_2[customer_data_2$Gender == "Female", ]
head(female_df)
```

```{r}
ggplot(data = male_df, mapping = aes(x=Annual.Income..k.., y=Spending.Score..1.100.)) +
  geom_point() +
  ggtitle("Income and Spending (Male)") +
  xlab("Annual Income ($1,000)") +
  ylab("Spending Score (1-100)") +
  theme_get()
```

```{r}
ggplot(data = female_df, mapping = aes(x=Annual.Income..k.., y=Spending.Score..1.100.)) +
  geom_point() +
  ggtitle("Income and Spending (Female)") +
  xlab("Annual Income ($1,000)") +
  ylab("Spending Score (1-100)") +
  theme_get()
```

### Plot Analysis Summary

!!!!!!!!  TODO  !!!!!!!!

-------------------------------------------------------------------

# Probability and Statistical Analysis

### Distribution

Use the dnorm function to verify whether or not the Age, Annual Income, and Spending Score variables follow the normal distribution
```{r}
age.range <- min(customer_data$Age):max(customer_data$Age)
age.mean <- mean(customer_data$Age)
age.sd <- sd(customer_data$Age)
age.dist <- dnorm(age.range, mean = age.mean, sd = age.sd)
age.df <- data.frame(Age = age.range, Density = age.dist)

ggplot(age.df, aes(x = Age, y = Density)) +
  geom_line() +
  ggtitle("Age Distribution")
```

```{r}
income.range <- min(customer_data$Annual.Income..k..):max(customer_data$Annual.Income..k..)
income.mean <- mean(customer_data$Annual.Income..k..)
income.sd <- sd(customer_data$Annual.Income..k..)
income.dist <- dnorm(income.range, mean = income.mean, sd = income.sd)
income.df <- data.frame(Income = income.range, Density = income.dist)

ggplot(income.df, aes(x = Income, y = Density)) +
  geom_line() +
  ggtitle("Annual Income Distribution") +
  xlab("Annual Income ($1,000's)")
```

```{r}
spending.range <- min(customer_data$Spending.Score..1.100.):max(customer_data$Spending.Score..1.100.)
spending.mean <- mean(customer_data$Spending.Score..1.100.)
spending.sd <- sd(customer_data$Spending.Score..1.100.)
spending.dist <- dnorm(spending.range, mean = spending.mean, sd = spending.sd)
spending.df <- data.frame(Spending = spending.range, Density = spending.dist)

ggplot(spending.df, aes(x = Spending, y = Density)) +
  geom_line() +
  ggtitle("Spending Score Distribution") +
  xlab("Spending Score (1-100)")
```

As we can see, all three of these variables (Age, Annual Income, and Spending Score) follow the normal distribution.

### Shapiro-Wilk Normality Test

Use the Shapiro-Wilk normality test to test for normality of each of the variables, and then each of the variables based on Gender.  

* What are the results? It they do not follow the normal distribution – which are closest?