---
title: "Chapter 13 - Iterating with purrr"
author: "Michael Beebe"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(purrr)
```

--------------------------------------------------------------------------------    

1. map
--
Create a list called myList. in this list store:  
 - A matrix of the numbers 1-10 in 2 rows  
 - The even numbers from 0-10  
 - The odd numbers from 1-9  
 - A matrix of the numbers from 1-9 in 3 rows  
```{r Create a list called myList. in this list store}
myList <- list(
  matrix(data = 1:10, nrow = 2),
  c(2, 4, 5, 6, 8, 10),
  c(1, 3, 5, 7, 9),
  matrix(data = 1:9, nrow = 3)
)

myList
```

Map the mean and sum functions to your list
```{r Map the mean and sum functions to your list}
map(myList, mean)
map(myList, sum)
```

Use the identical function to determine if map and lapply are the same when applied to myList
```{r Use the identical function to determine if map and lapply are the same when applied to myList}
lapply(myList, mean)
lapply(myList, sum)
```

2. map with Specified Types
--
Using your list - myList - find the following:  
 - Use map_int to find number of rows or length of each list element  
```{r Use map_int to find number of rows or length of each list element}
map_int(myList, length)
```

 - Use map_dbl to find the standard deviation  
```{r Use map_dbl to find the standard deviation}
map_dbl(myList, sd)
```
 
 - Use map_chr to find the class of each element in your list  
```{r Use map_chr to find the class of each element in your list}
map_chr(myList, typeof)
```
 
 - Use map_lgl to determine if the number of rows in each element of your list is < 3  
```{r Use map_lgl to determine if the number of rows in each element of your list is < 3}
check <- function(x) is.list(nrow(x)) < 3
map_lgl(myList, check)
```
 
 - Create the data frame and list on page 184 of your text. Use map_df to call the function with the list of lengths  
```{r Create the data frame and list on page 184 of your text}
buildDF <- function(x) {
  data.frame(A=1:x, B=x:1)
}

listOfLengths <- list(3, 4, 1, 5)

listOfLengths %>%
  map(buildDF)
```

3. Iterating over a data.frame
--
Use the mtcars built in data set  
Find the mean of all of the numeric data in the dataset
```{r Find the mean of all of the numeric data in the dataset}
means <-
  mtcars %>%
    map(mean)

means
```

Find the standard deviation of all of the numeric data in the dataset
```{r Find the standard deviation of all of the numeric data in the dataset}
stdvs <-
  mtcars %>%
    map(sd)

stdvs
```


