---
title: 'Lab Assignment: Chapter 14 - Data Reshaping'
author: "Michael Beebe"
output:
  html_document: default
  word_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Dependencies}
library(stringr)
library(plyr)
library(reshape2)
library(ggplot2)
library(scales)
```

1. cbind and rbind
--  

Create 3 vectors: start, finish, and numeric  

* The start vector has the first 10 capital letters of the alphabet  
* The finish vector has the last 10 capital letters of the alphabet  
* The numeric vector has the numbers 1-10
```{r Create 3 vectors: start, finish, and numeric }
start <- c("A", "B", "C", "D", "E", "F", "G", "H", "I", "J")
finish <- c("Q", "R", "S", "T", "U", "V", "W", "X", "Y", "Z")
numeric <- c(1:10)
```

Use cbind to bind these vectors into table1 by column
```{r Use cbind to bind these vectors into table1 by column}
table1 <- cbind(start, finish, numeric)
table1
```

Use rbind to bind these vectors into table2 by row
```{r Use rbind to bind these vectors into table2 by row}
table2 <- rbind(start, finish, numeric)
table2
```

Create 2 CSV files - data1.csv and data2.csv  

* data1.csvshould have the column titles "Course" and "Title"   
-- List 7 courses and titles in the Data Analytics Degree  
* data2.csv should have the column titles "Credits" and "Semester"  
-- List the credit hours and the semester you should take the courses in data1.csv  
```{r Create 2 CSV files - data1.csv and data2.csv}
# DFs THAT WE WILL USE TO MAKE THE CSVs
data1 <- data.frame(
  Course = c("ITCS 1010", "ITDB 1300", "ITDB 1400", "ITDB 1405", "ITCS 1880", "ITON 1205", "ITCS 1870"),
  Title = c("Programming Logic", "Intro to Database Theory", "Intro to SQL", "Oracle PL/SQL Programming", "R Programming", "Network+ and Networking Essentials", "Python Programming")
)

data2 <- data.frame(
  Credits = c(3, 3, 2, 2, 3, 3, 3),
  Semester = c(2, 2, 2, 2, 3, 3, 4)
)


# MAKING CSVs FROM DF
data1.csv <- write.csv(
  x = data1,
  file = "./Part1CSVs/data1.csv"
)

data2.csv <- write.csv(
  x = data2,
  file = "./Part1CSVs/data2.csv"
)
```

Import and use cbind to bind these 2 datasets together.
```{r Import and use cbind to bind these 2 datasets together.}
data3.csv <- write.csv(
  x = cbind(data1, data2),
  file = "./Part1CSVs/data3.csv"
)
```

2. Joins
--
Unzip the US_Foreign_Aid.zip file into a directory on your computer
```{r}
foreign_aid <- unzip("../../Data/US_Foreign_Aid.zip", exdir = "./US_Foreign_Aid")
foreign_aid
```
```{r}
aid_data <- dir("US_Foreign_Aid/", pattern = "\\.csv")

for (i in aid_data) {
  us_foreign_aid <- str_sub(string = i, start=12, end=18)
  temp <- read.table(file = file.path("US_Foreign_Aid/", i),
                     header = TRUE, sep = ",", stringsAsFactors = FALSE)
  assign(x=us_foreign_aid, value = temp)
}
```

Merge the data from the 80s and 90s and display to the console
```{r Merge the data from the 80s and 90s and display to the console}
aid80s90s <- merge(
  x = Aid_80s,
  y = Aid_90s,
  by.x = c("Country.Name", "Program.Name"),
  by.y = c("Country.Name", "Program.Name")
)

head(aid80s90s)
```

Merge the data from the 90s and 00s and display to the console
```{r Merge the data from the 90s and 00s and display to the console}
aid90s00s <- merge(
  x = Aid_90s,
  y = Aid_00s,
  by.x = c("Country.Name", "Program.Name"),
  by.y = c("Country.Name", "Program.Name")
)

head(aid90s00s)
```

Use the plyr package and join function to do the same for both sets of data  
```{r Use the plyr package and join function to do the same for both sets of data}
aid80s90s_join <- join(
  x = Aid_80s,
  y = Aid_90s,
  by = c("Country.Name", "Program.Name")
)

aid90s00s_join <- join(
  x = Aid_90s,
  y = Aid_00s,
  by = c("Country.Name", "Program.Name")
)
```

Combine the 8 foreign aid data frames into 1 data frame by putting them in a list and joining toghether  

* Test your list with different types of subscripts
```{r Combine the 8 foreign aid data frames into 1}
df_names <-str_sub(string = aid_data, start = 12, end = 18)
df_list <- vector("list", length(df_names))
names(df_list) <- df_names
for (i in df_names) {
  df_list[[i]] <- eval(parse(text = i))
}

all_aid <- Reduce(function(...) {
  join(..., by = c("Country.Name", "Program.Name"))
}, df_list)

dim(all_aid)

```

3. reshape2
--

Use melt so that each row represents a Country-Program-Year entry with Dollars stored in a column  
```{r Use melt so that each row represents a Country-Program-Year entry with Dollars stored in a column}
aid_melt <- melt(
  data = all_aid,
  id.vars = c("Country.Name", "Program.Name"),
  variable.name = "Year",
  value.name = "Dollars"
)

aid_melt$Year <- as.numeric(str_sub(aid_melt$Year, start = 3, end = 6))

tail(aid_melt)
```

Aggregate and plot the data
```{r Aggregate and plot the data}
melt_agg <- aggregate(
  Dollars ~ Program.Name + Year,
  data = aid_melt,
  FUN = sum,
  na.rm = TRUE
)

melt_agg$Program.Name <- str_sub(melt_agg$Program.Name, start = 1, end = 10)

ggplot(melt_agg, aes(x=Year, y=Dollars)) +
  geom_line(aes(group=Program.Name))
```

